---
title: "Dataprep_1.1"
output: html_document
date: "2023-11-22"
---
```{r setup}

library(haven)
library(tidyverse)
library(tidyr)
library(mokken)
library(OpenMx)
library(ggplot2)
library(lavaan)
library(semTools)
library(semPlot)
library(psych)
library(GPArotation)
library(corrplot)
library(stats)
library(dplyr)
library(usethis)
library(gitcreds)
library(gh)
library(repro)

```


```{r}
knitr::opts_chunk$set(echo = TRUE)

## Wave 1: enread data 
sharew1_rel8_0_0_cv_r <- read_dta("sharew1_rel8-0-0_ALL_datasets_stata/sharew1_rel8-0-0_cv_r.dta")
sharew1_rel8_0_0_gv_health <- read_dta("sharew1_rel8-0-0_ALL_datasets_stata/sharew1_rel8-0-0_gv_health.dta")
sharew1_rel8_0_0_dropoff <- read_dta("sharew1_rel8-0-0_ALL_datasets_stata/sharew1_rel8-0-0_dropoff.dta")


#####
attach(sharew1_rel8_0_0_cv_r)
cvr1_items <-data.frame(mergeid,country,age2004)

attach(sharew1_rel8_0_0_dropoff)
dr_casp_items<- data.frame(mergeid,q2_a,q2_b,q2_c,q2_d,q2_e,q2_f,q2_g,q2_h,
q2_i,q2_j,q2_k,q2_l) ## select items


cvr_all <- full_join(cvr1_items, dr_casp_items,by = "mergeid")
 
#show(cvr_all) #commented out for knitting 
dim(cvr_all)  
names(cvr_all)
detach()
```
In the following chunk of code I selected the respective countries of the data.
I then also got rid of the NAs in the data. And I made sure, that the variables 
are numerics in the data frame.
# country als dichotom
```{r}
cvr_notall <- subset(cvr_all, country %in% c(12, 13), 
                     select = c(country,age2004, q2_a, q2_b, q2_c, q2_d, q2_e, 
                                q2_f, q2_g, q2_h, q2_i, q2_j, q2_k, q2_l))
cvr_notall$country <- as.numeric(cvr_notall$country)

cvr_notall$country[cvr_notall$country == 12] <- 1
cvr_notall$country[cvr_notall$country == 13] <- 2

cvr_notall[cvr_notall %in% c(-1, -2, -3, -4, -5, -7, -9, -10, -11, -12, -13, -14, -15, -16)] <- NA
cvr_notall <- na.omit(cvr_notall)

```

# Descriptives of the data used
```{r}
describe(cvr_notall)

cor_matrix <- cor(cvr_notall)
# Display the correlation matrix
print(cor_matrix)

# Calculate item variances
item_variances <- apply(cvr_notall[, c("q2_a", "q2_b", "q2_c", "q2_d", "q2_e", "q2_f", "q2_g", "q2_h", "q2_i", "q2_j", "q2_k", "q2_l")], 2, var)

# Print the item variances
print(item_variances)
```

# Z-scale the variables in the dataframe
## Troubleshoot attempt
```{r}
cvr_notall_scaled <- as.data.frame(scale(cvr_notall))

```

```{r}
cvr_long <- pivot_longer(cvr_notall, cols = -c(1:3), 
                        names_to = c('.value', 'wave'), 
                        names_pattern = '(.*)(\\d+)')

```


```{r}
## Recode negatively worded items
names(cvr_notall)

cvr_notall$q2_d <- 5 - cvr_notall$q2_d
cvr_notall$q2_g <- 5 - cvr_notall$q2_g
cvr_notall$q2_h <- 5 - cvr_notall$q2_h
cvr_notall$q2_i <- 5 - cvr_notall$q2_i
cvr_notall$q2_j <- 5 - cvr_notall$q2_j
cvr_notall$q2_k <- 5 - cvr_notall$q2_k
cvr_notall$q2_l <- 5 - cvr_notall$q2_l

## Standardize age
cvr_notall$age2004 <- (cvr_notall$age2004 - mean(cvr_notall$age2004))/sd(cvr_notall$age2004)

```
```{r}
df4_1 <- cvr_notall %>% 
        rename(
        "age" = "age2004"
               ) 
```

```{r}
mxdata1 <- mxData(observed=df4_1,type="raw")
manVars <- colnames(df4_1[,-c(1,2)])
nv <- length(manVars)

```

# CFA
```{r}
#measurement-models CTT
casp_g.tau.kon.mod <- 'control  =~ q2_a + q2_b + q2_c
                       autonomy  =~ q2_d + q2_e + q2_f
                       pleasure  =~ q2_g + q2_h + q2_i
                       self_real  =~ q2_j + q2_k + q2_l
                     '
casp_g.tau.kon.fit<- cfa(casp_g.tau.kon.mod, dr_casp_items,
                        meanstructure = TRUE,estimator="MLR")
summary(casp_g.tau.kon.fit,fit.measures=TRUE,standardized=TRUE)

# Loadings set to be equal
casp_g.ess.tau.q.mod <- 'control  =~ q2_a + 1*q2_b + 1*q2_c
                         autonomy  =~ q2_d + 1*q2_e + 1*q2_f
                         pleasure  =~ q2_g + 1*q2_h + 1*q2_i
                         self_real  =~ q2_j + 1*q2_k + 1*q2_l
                        '
casp_g.ess.tau.q.fit<- cfa(casp_g.ess.tau.q.mod, dr_casp_items,
                           meanstructure = TRUE,estimator="MLR")
summary(casp_g.ess.tau.q.fit,fit.measures=TRUE,standardized=TRUE)

lavTestLRT(casp_g.tau.kon.fit, casp_g.ess.tau.q.fit)

```

# Secondorder CFA - the model
# latent variable definitions
```{r}
casp_g.tau.kon.mod <- 'control  =~ q2_a + q2_b + q2_c
                       autonomy  =~ q2_d + q2_e + q2_f
                       pleasure  =~ q2_g + q2_h + q2_i
                       self_real  =~ q2_j + q2_k + q2_l
                     '
#CFA
casp_cfa_mod <- cfa(casp_g.tau.kon.mod, dr_casp_items, estimator="MLR") 
summary(casp_cfa_mod,fit.measures=TRUE,standardized=TRUE)
mod_i_m2a <- modificationIndices(casp_cfa_mod, sort=TRUE, minimum.value = 3)
view(mod_i_m2a)
```

# residualcovariancematrix
```{r}
resid(casp_cfa_mod)
```

The SemPaths I just did for cosmetic reasons. I always like to be able to view 
the model. And that way I could take a look at it, consisely, without any 
tables.

# SEM model
```{r}
sem_casp_cfa_mod <- sem(casp_g.tau.kon.mod, data = dr_casp_items, estimator= "mlr")
summary(sem_casp_cfa_mod, standardized=TRUE)
fitmeasures(sem_casp_cfa_mod, c("cfi","rmsea","tli","srmr"))
semPaths(sem_casp_cfa_mod, what="paths",whatLabels= "stand", rotation=1)

```

### MI
# configural invariance
```{r}
fit1_config <- cfa(casp_g.tau.kon.mod, data = cvr_notall, estimator= "MLR",
                       group = "country")
summary(fit1_config, fit.measures = TRUE, standardized = TRUE)
```

# Metric model
```{r}
fit2_metric <- cfa(casp_g.tau.kon.mod, data = cvr_notall, estimator = "MLR",
                       group = "country", group.equal = "loadings")
summary(fit2_metric, fit.measures = TRUE, standardized = TRUE)

summary(compareFit(fit1_config, fit2_metric))
```

# Scalar model
```{r}
fit3_scalar <- cfa(casp_g.tau.kon.mod, data = cvr_notall, estimator = "MLR",
                       group = "country", group.equal = c("loadings", "intercepts"))
summary(fit3_scalar, fit.measures = TRUE, standardized = TRUE)

summary(compareFit(fit2_metric, fit3_scalar))
```