---
title: "MNLFA_age_VK"
author: "Valentin Kriegmair"
date: "2023-06-25"
output:
  html_document:
    number_sections: true
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Primary Libraries
```{r}
library(haven)
library(OpenMx)
library(tidyverse)
library(psych)
library(semPlot)
library(lavaan)
library(corrplot)
```

# Read and Prepare Data
```{r}
# read
setwd("C:/Users/admin/Desktop/MPI/Share/ShareSEM/ShareSEM")
df_ac7 <- read_dta("sharew7_rel8-0-0_ALL_datasets_stata/sharew7_rel8-0-0_ac.dta")
demo7 <- read_dta("sharew7_rel8-0-0_ALL_datasets_stata/sharew7_rel8-0-0_dn.dta")
demo7items <- demo7 %>% dplyr::select("mergeid", "dn003_", "dn042_") 
View(demo7)
# prepare Data: 
df0 <- df_ac7 %>% 
  mutate(
    # name CASP-12 Items as variables according to factor structure.
    cC1 = ac014_,
    cC2 = ac015_,
    cC3 = ac016_,
    cA1 = ac017_,
    cA2 = ac018_,
    cA3 = ac019_,
    cP1 = ac020_,
    cP2 = ac021_,
    cP3 = ac022_,
    cS1 = ac023_,
    cS2 = ac024_,
    cS3 = ac025_
  )

# merge demographic variables to items:
df1 <- df0 %>% 
  full_join(
    demo7items, by="mergeid"
  ) %>% 
  rename(
    age = dn003_,
    gender = dn042_
  ) %>% 
  dplyr::select(gender, age, matches("^cC|^cA|^cP|^cS"))

# Storing the names of variables (items) as `manVars`:
manVars <- colnames(df1[,-c(1:2)])

# Storing the number of variables (items) as `nv`:
nv <- length(manVars)
```

## Missing Values

```{r missing}
# Define Missing Codes. Refer to the SHARE release guide for more detail:
missing_codes <- c(-1, -2, -3, -7, -9, -10, -11, -12, -14, -15, -16)

df2 <- df1 %>%
  mutate_all(~ifelse(. %in% missing_codes, NA, .))

# Check number of subjects with NAs on all CASP-12:
n_missings <- 
  sum(apply(df2[manVars], 1, function(row) all(is.na(row))))

# Remove 7623 rows with missing values on all CASP-12 variables:
df3 <- df2 %>% 
  filter(rowSums(is.na(.[manVars])) < length(manVars)) %>% 
# Remove missings on moderators gender and `age` (year of birth) and bfi items
  filter(complete.cases(gender, age))
```


## Recode Items
```{r}
df4 <- df3 %>% 
  # Recode negative items of the Casp-12 questionnaire:
  mutate(
    cA1 = 5 - cA1,
    cP1 = 5 - cP1,
    cP2 = 5 - cP2,
    cP3 = 5 - cP3,
    cS1 = 5 - cS1,
    cS2 = 5 - cS2,
    cS3 = 5 - cS3,
  # Recode gender variable to `male` with 1 indicating male and 0 female:
    male = 2 - gender,
    age = 2017 - age
  ) %>% 
  # Exclude constructor variables:
  dplyr::select(-gender) %>% 
  # Order variables:
  dplyr::select(male, age, everything())

# Standardize all variables
df_z <- df4 %>% 
  mutate_if(names(.) != "male", scale)
```

## Descriptives
```{r}
hist(df4$age)
#hist(df4$bfO)

df_long <- df4 %>% 
  dplyr::select(starts_with("c"))
  pivot_longer(cols = everything(),
    names_to = "casp_12",
    values_to = "value")

  
  # Create the plot
ggplot(df_long, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ casp_12, scales = "free_y") 
```

```{r}
cor <- cor(df4, use= "pairwise.complete.obs")

corrplot(cor, method = "color")
```

## Fit configural model
```{r}
model <- "
  cntrl =~ NA*cC1 + cC2 + cC3
  aut =~ NA*cA1 + cA2 + cA3
  pleas =~ NA*cP1 + cP2 + cP3
  selfre =~ NA*cS1 + cS2 + cS3

  cntrl ~~ 1 * cntrl
  aut ~~ 1 * aut
  pleas ~~ 1 * pleas
  selfre ~~ 1 * selfre
"

fit <- cfa(model, data = df_z, estimator = "MLR", group = "male")

summary(fit, fit.measures = TRUE)

# semPaths(fit, "est", curveAdjacent = TRUE)

# modindices(fit, sort = TRUE, maximum.number = 20)


fit1 <- cfa(model, data = df_z, estimator = "MLR")

summary(fit1, fit.measures = TRUE)

semPaths(fit1, "est", curveAdjacent = TRUE)

fit_indices <- fitMeasures(fit1)

fit_table <- tibble(FitIndex = names(fit_indices), Value = fit_indices)
desired_indices <- c("rmsea", "rmr", "srmr", 
                     "nfi", "nnfi", "cfi",
                     "pnfi")
fit_table_subset <- fit_table[fit_table$FitIndex %in% desired_indices, ]
fit_table_subset$Value <- round(as.numeric(fit_table_subset$Value), 3)

# Print the table
# fit_table_subset %>%
  #  kable("html", caption = "Fit Indices")

# modindices(fit, sort = TRUE, maximum.number = 20)
```

```{r}
# Saving data frame as mx object:
mxdf1 <- mxData(observed = df_z, type = "raw")
```

## Assessing Measurement Invariance
- 1. comparing fit of unconstrained configural model with the fit of metric model
- 2. comparing fit of metric model with the fit of scalar model
## Configural Model:
### Base Matrices
```{r}
# Item Intercepts
# Baseline intercepts:
matT0 <- mxMatrix(type="Full", nrow = 1, ncol=nv,
                  free=TRUE, values = 1, name="matT0")

# Effect of the background variable `age` on the intercepts (freely estimated):
matB1 <-  mxMatrix(type="Full", nrow = 1, ncol=nv,
                   free=TRUE, values = 0, name="matB1")

# Item Loadings on respective Factors
# Baseline factor loadings:
matL0 <- mxMatrix(type="Full", nrow=nv, ncol = 4,
                   free=c(rep(c(T, F, F, F),3), 
                          rep(c(F, T, F, F),3),
                          rep(c(F, F, T, F),3),
                          rep(c(F, F, F, T),3)),
                   values=c(rep(c(1, 0, 0, 0),3), 
                            rep(c(0, 1, 0, 0),3),
                            rep(c(0, 0, 1, 0),3),
                            rep(c(0, 0, 0, 1),3)),
                   byrow = T,
                   name="matL0")

# Effect of `age` on the factor loadings (freely estimated):
matC1 <- mxMatrix(type = "Full",
                  nrow = nv,
                  ncol = 4,
                  free= c(rep(c(T, F, F, F),3), 
                          rep(c(F, T, F, F),3),
                          rep(c(F, F, T, F),3),
                          rep(c(F, F, F, T),3)),
                  values = 0,
                  byrow = T,
                  name = "matC1"
                  )

# Residual Variances:
# baseline residual variances
matE0 <- mxMatrix( type = "Diag", nrow = nv,
                   ncol = nv,
                   free = T ,
                   values = 1,
                   name = "matE0")

# Effect of `age` on residual variances
matD1 <- mxMatrix( type = "Diag", nrow = nv,
                   ncol = nv,
                   free = T ,
                   values = 0,
                   name = "matD1")

# Baseline common-factor variances and the baseline correlation:
matP0 <- mxMatrix(type = "Symm", nrow = 4,
                  ncol = 4,
                  free = c(F, T, T, T,
                           T, F, T, T,
                           T, T, F, T,
                           T, T, T, F),
                  values = 1,
                  name = "matP0"
                  )

# Effect of `age` on factor (variances and) correlations:
matH1 <- mxMatrix(type = "Symm", nrow = 4,
                  ncol = 4,
                  free = c(F, T, T, T,
                           T, F, T, T,
                           T, T, F, T,
                           T, T, T, F),
                  values = 0,
                  name = "matH1")
# note. Only effects on correlations are estimated,
# effects on variances are set to 0 

# Baseline factor means (fixed to 0):
matA0 <- mxMatrix(type = "Full", nrow = 4, ncol = 1, 
                  free = F, values = 0, name = "matA0")

# Effect of `age` on factor means (fixed to 0):
matG1 <- mxMatrix(type = "Full", nrow = 4, ncol = 1, 
                  free = F, values = 0, name = "matG1")

### Background Variables
# observed scores of `age`:
matV1 <- mxMatrix(type = "Full", nrow = 1,
                  ncol = 1,
                  free = F,
                  labels = "data.age",
                  name = "age")
```

### Algebraic Matrices
```{r}
### Parameters predicted by Background Variables
# Identity Matrices for conversion:
matIa <- mxMatrix(type="Diag", nrow=4, ncol=4, 
                  free=FALSE,
                  values=1, 
                  name="matIa")

matIb <- mxMatrix(type="Full", nrow=4, ncol=4, 
                  free=FALSE, 
                  values=c(0,1,1,1,
                           1,0,1,1,
                           1,1,0,1,
                           1,1,1,0),
                  name="matIb")

# Intercepts:
matT <- mxAlgebra(expression=matT0+matB1*age, 
                  name="matT")
# general intercept, male variable, age variable, named matT

# Loadings:
matL <- mxAlgebra(expression=matL0+matC1*age, 
                  name="matL")


# Latent Variance:
matE <- mxAlgebra(expression=matE0*exp(matD1*age), 
                  name="matE")

# Latent Intercept:
matA <- mxAlgebra(expression=matA0+matG1*age, 
                  name="matA")

## Specify algebra for covariance matrix of factors (transformed to ensure positive definite matrices)

# Factor var-cov matrix predicted by age
matVar <- mxAlgebra(expression=(matP0*exp(matH1*age)), 
                    name="matVar")

# Factor corr predicted by age
matR <- mxAlgebra(expression=(exp(2*(matP0+matH1*age))-1)/
                    (exp(2*(matP0+matH1*age))+1), 
                  name="matR")

matCov <- mxAlgebra(expression=(matIa*sqrt(matVar))%*%matR%*%(matIa*sqrt(matVar)), 
                    name="matCov")
matP <- mxAlgebra(expression=matIa*matVar+matIb*matCov, 
                  name="matP")

## Specify model-implied matrices
matC <- mxAlgebra(expression=matL%*%matP%*%t(matL)+matE, 
                  name="matC") 
matM <- mxAlgebra(expression=matT+t(matL%*%matA), 
                  name="matM") 
```

### Configural Model Fit
```{r}
expF <- mxExpectationNormal(covariance="matC", 
                            means="matM",
                            dimnames=manVars)
fitF <- mxFitFunctionML() 

modConfig <- mxModel(model="Configural", 
                     matT, matT0, matB1, 
                     matL, matL0, matC1, 
                     matE, matE0, matD1, 
                     matP, matP0, matH1, 
                     matA, matA0, matG1, 
                     matIa, matIb, matV1,
                     matVar, matR, matCov, matM, matC, 
                     expF, fitF, mxdf1)

fitConfig <- mxRun(modConfig)

summary(fitConfig)
```

### Metric Model
fix the effect of age on the item loadings to 0.
```{r}
# respecify only matrix with item loadings (fixed to 0)
matC1 <- mxMatrix(type="Full", nrow=nv, ncol=4,
                  free=FALSE,
                  values=0,
                  name="matC1")

## Make mxModel object and run the model
modMetric <- mxModel(model="Metric", 
                     matT, matT0, matB1,
                     matL, matL0, matC1,  
                     matE, matE0, matD1, 
                     matP, matP0, matH1, 
                     matA, matA0, matG1, 
                     matIa, matIb, matV1,
                     matVar, matR, matCov, matM, matC, 
                     expF, fitF, mxdf1)
fitMetric <- mxRun(modMetric)
summary(fitMetric)

miTest <- mxCompare(fitConfig, fitMetric)
miTest
fitConfig$matC1
fitMetric$matH1
```

### Scalar Model
```{r}
## Specify matrices scalar model
# here all fixed to 0, as there is no moderator here.

# Fix effects of age on item intercepts to 0
matB1 <- mxMatrix(type="Full", nrow=1, ncol=nv,
                  free=FALSE,
                  values=0,
                  name="matB1")

# Freely estimate effect of age on factor variances
matH1 <- mxMatrix(type="Symm", nrow=4, ncol=4,
                  free=TRUE, 
                  values=0,
                  name="matH1")

# Freely estimate effect of age on factor means
matG1 <- mxMatrix(type="Full", nrow=4, ncol=1,
                  free=TRUE, 
                  values=0,
                  name="matG1")


## Make mxModel object and run the model
modScalar <- mxModel(model="Scalar", 
                     matT, matT0, matB1,
                     matL, matL0, matC1,  
                     matE, matE0, matD1, 
                     matP, matP0, matH1, 
                     matA, matA0, matG1, 
                     matIa, matIb, matV1,
                     matVar, matR, matCov, matM, matC, 
                     expF, fitF, mxdf1)
fitScalar <- mxRun(modScalar)
summary(fitScalar)

miTest <- mxCompare(fitMetric, fitScalar)
miTest

fitScalar$matG1
```
the scalar model fits the data (significantly) worse than the and configural (and metric) model, suggesting that full scalar invariance does not hold.

# Test for Partial Invariance
## Unconstrained Model for each Item
```{r}
## Run unconstrained model for each indicator
fitAbo <- list()
# all but one: assume all items to be invariant except one, then next and so on
# …

for (i in 1:12){
  freeparT <- matrix(FALSE, nrow=1, ncol=12)
  freeparT[i] <- TRUE
  freeparL <- matrix(FALSE, nrow=12, ncol=4)
  freeparL[i,ifelse(i<4,1,ifelse(i<7,2,ifelse(i<10,3,4)))] <- TRUE
  matB1 <- mxMatrix(type="Full", nrow=1, ncol=nv,
                    free=freeparT,
                    values=0,
                    name="matB1")
  matC1 <- mxMatrix(type="Full", nrow=nv, ncol=4,
                    free=freeparL,
                    values=0,
                    byrow=TRUE,
                    name="matC1")
  modAbo <- mxModel(model=paste0("All_but_", i), 
                   matT, matT0, matB1,
                   matL, matL0, matC1,
                   matE, matE0, matD1,
                   matP, matP0, matH1,
                   matA, matA0, matG1,
                   matIa, matIb, matV1,
                   matVar, matR, matCov, matM, matC, 
                   expF, fitF, mxdf1)
  fitAbo[[i]] <- mxRun(modAbo)
}

## Compare constrained model with all unconstrained models
anchorTest <- mxCompare(fitAbo, fitScalar)
anchorOut <- data.frame(Name=manVars, 
                    X2=anchorTest$diffLL[seq(2,24,2)],
                    df=anchorTest$diffdf[seq(2,24,2)],
                    p=anchorTest$p[seq(2,24,2)])
# anchorOut
# this puts out
```

## Select Anchor Items
```{r}
## Select two indicators per factor with smallest X2 as anchor
## number of manifest items. rule of thumb: 20% per indicator but at least 2.
# Look at the all but one models for each factor ordered by X2 value:
# For Control:
anchorOut[order(anchorOut$X2[1:3]),]
# For Autonomy:
anchorOut[3 + order(anchorOut$X2[4:6]),]
# For Pleasure:
anchorOut[6 + order(anchorOut$X2[7:9]),]
# For Selfrealization:
anchorOut[9 + order(anchorOut$X2[10:12]),]
anchorOut
# Get the two items with the smallest X2 for each factor.
anchors1 <- as.numeric(rownames(head(anchorOut[order(anchorOut$X2[1:3]), ], 2)))
anchors2 <- as.numeric(rownames(head(anchorOut[3 + order(anchorOut$X2[4:6]),], 2)))
anchors3 <-  as.numeric(rownames(head(anchorOut[6 + order(anchorOut$X2[7:9]),], 2)))
anchors4 <- as.numeric(rownames(head(anchorOut[9 + order(anchorOut$X2[10:12]),], 2)))
```

## Anchors-Plus-One Models
Sanity check free and fixed values.
```{r}
# vector that indicates free values for item intercepts:
freeparT <- matrix(data=TRUE, nrow=1, ncol=12)
freeparT[1,c(anchors1,anchors2, anchors3, anchors3)] <- FALSE

# vector that indicates free values for item loadings: 
freeparL <- matrix(rep(as.logical(diag(4)), each = 3), nrow = nv)
freeparL[anchors1,1] <- F
freeparL[anchors2,2] <- F
freeparL[anchors3,3] <- F
freeparL[anchors4,4] <- F

freeparT
freeparL
```

```{r}
## Specify matrices for unconstrained model
# get estimates for 
# - `bf0` effects on factor correlations (variances fixed) (H1)
# -  base / intercepts of factor correlations (variances fixed) (P0)
# from the scalar model as starting values.
scalarH1 <- fitScalar$matH1$values
scalarP0 <- fitScalar$matP0$values
scalarP0
# get estimates for 
# - `bfO` effects on indicator loadings (C1) only for free parameters 
# (others set to 0)
# from the configural models as starting values
configC1 <- fitConfig$matC1$values
startingC1 <- matrix(0, nrow = 12, ncol = 4)
startingC1[freeparL] <- configC1[freeparL]

# configL0 <- fitConfig$matL0$values
# scalarT0 <- fitScalar$matT0$values
matT0 <- mxMatrix(type="Full", nrow = 1, ncol=nv,
                  free=TRUE, values = 0, name="matT0")
matB1 <- mxMatrix(type="Full", nrow=1, ncol=nv,
                  free=freeparT,
                  values=0,
                  name="matB1")
matL0 <- mxMatrix(type="Full", nrow=nv, ncol = 4,
                   free=c(rep(c(T, F, F, F),3), 
                          rep(c(F, T, F, F),3),
                          rep(c(F, F, T, F),3),
                          rep(c(F, F, F, T),3)),
                   values=0,
                   byrow = T,
                   name="matL0")
matC1 <- mxMatrix(type="Full", nrow=nv, ncol=4,
                  free=freeparL,
                  values=startingC1,
                  name="matC1")
matP0 <- mxMatrix(type = "Symm", nrow = 4,
                  ncol = 4,
                  free = c(F, T, T, T,
                           T, F, T, T,
                           T, T, F, T,
                           T, T, T, F),
                  values = scalarP0,
                  name = "matP0"
                  )
matH1 <- mxMatrix(type="Symm", nrow=4, ncol=4,
                  free=TRUE, 
                  values=scalarH1,
                  name="matH1")
startingC1
```

```{r}
# Unconstrained model with only "anchor items" allowed to vary as function of
# `age`
modAnchors <- mxModel(model="AnchorsOnly", 
                      matT, matT0, matB1,
                      matL, matL0, matC1,
                      matE, matE0, matD1,
                      matP, matP0, matH1,
                      matA, matA0, matG1,
                      matIa, matIb, matV1,
                      matVar, matR, matCov, matM, matC, 
                      expF, fitF, mxdf1)
fitAnchors <- mxRun(modAnchors)
summary(fitAnchors)

mxCompare(fitConfig, fitAnchors)
```

```{r}
## Run constrained model for each indicator (except the anchors)
testIn <- c(1:12)[-c(anchors1, anchors2, anchors3, anchors4)]

# Initialising a list to store the results of model runs.
fitApo <- list()

anchorsC1 <- fitAnchors$matC1$values
anchorsB1 <- fitAnchors$matB1$values

anchorsH1 <- fitAnchors$matH1$values
anchorsP0 <- fitAnchors$matP0$values

matP0 <- mxMatrix(type = "Symm", nrow = 4,
                  ncol = 4,
                  free = c(F, T, T, T,
                           T, F, T, T,
                           T, T, F, T,
                           T, T, T, F),
                  values = anchorsP0,
                  name = "matP0"
                  )

matH1 <- mxMatrix(type="Symm", nrow=4, ncol=4,
                  free=TRUE, 
                  values=anchorsH1,
                  name="matH1")
```

```{r}
for (i in testIn){
  freeparTa <- freeparT # Copying the original `freeparT` matrix
  freeparLa <- freeparL # Copying the original `freeparL` matrix
  freeparTa[1,i] <- F # Setting the 'i'th column of the first row in 'freeparTa' to FALSE
  # if (freeparTa[1,i] == F){ anchorsB1[1,i] <- 0
  # }
  # Defining the column index based on the value of 'i':
  indexLa <- ifelse(i<4,1,ifelse(i<7,2,ifelse(i<10,3,4)))
  # fixing parameter depending on index:
  freeparLa[i,indexLa] <- F
  # Checking if the corresponding element in 'freeparLa' is FALSE:
  # if(freeparLa[i, indexLa] == F){
    # If TRUE, setting the corresponding starting value of matC1 to 0.
    # startingC1[i, indexLa] <- 0
  # }
  matB1 <- mxMatrix(type="Full", nrow=1, ncol=nv,
                    free=freeparTa,
                    values=0,
                    name="matB1")
  matC1 <- mxMatrix(type="Full", nrow=nv, ncol=4,
                    free=freeparLa,
                    values=0, #starting values from 
                    name="matC1")
  modApo <- mxModel(model=paste0("Anchors_plus_", i), 
                    matT, matT0, matB1,
                    matL, matL0, matC1,
                    matE, matE0, matD1,
                    matP, matP0, matH1,
                    matA, matA0, matG1,
                    matIa, matIb, matV1, 
                    matVar, matR, matCov, matM, matC, 
                    expF, fitF, mxdf1)
  fitApo[[i]] <- mxRun(modApo)
}


# retrieve the BIC of the configural model
config_bic <- summary(fitConfig)$BIC
apo_bics <- sapply(fitApo[c(1, 4, 9, 10)], function(x) summary(x)$BIC)

config_bic - apo_bics

result <- sapply(apo_bics, function(x) config_bic - x)
result

summary(fitConfig)$BIC - summary(fitApo[[1]])$BIC
fitConfig$matB1

```

```{r}
# Specify only model with index 4 with different starting values:
freeparTa <- freeparT 
freeparLa <- freeparL 


# Set wanted index to 4:
i <- 1
# Set the i-th column of the first row in 'freeparTa' to FALSE
freeparTa[1, i] <- F 

# Define the column index based on the value of 'i':
indexLa <- ifelse(i<4, 1, ifelse(i<7, 2, ifelse(i<10, 3, 4)))

freeparLa[i, indexLa] <- F

#Starting values:
anchorsB1 <- fitApo[[1]]$matB1$values
anchorsT0 <- fitApo[[1]]$matT0$values

# Adjust starting values matrix to have 0 on respective fixed elements:
if (freeparTa[1,i] == F){
      anchorsB1[1,i] <- 0
}

matT0 <- mxMatrix(type="Full", nrow = 1, ncol=nv,
                  free=TRUE, values = anchorsT0, name="matT0")

matB1 <- mxMatrix(type="Full", nrow=1, ncol=nv,
                  free=freeparTa,
                  values=anchorsB1,
                  name="matB1")

matC1 <- mxMatrix(type="Full", nrow=nv, ncol=4,
                  free=freeparLa,
                  values=0,
                  name="matC1")

modApo <- mxModel(model=paste0("Anchors_plus_", i), 
                  matT, matT0, matB1,
                  matL, matL0, matC1,
                  matE, matE0, matD1,
                  matP, matP0, matH1,
                  matA, matA0, matG1,
                  matIa, matIb, matV1, 
                  matVar, matR, matCov, matM, matC, 
                  expF, fitF, mxdf1)

fitApo1 <- mxRun(modApo)
# fitApo1 <- mxTryHard(fitApo4)
# add the model to the preexisting list:
fitApo[[1]] <- fitApo1
```


## Likelihood-Ratio Tests
```{r}
piTest <- mxCompare(fitAnchors, fitApo)
piTest
piOut <- data.frame(Name=paste0("Indicator",testIn),
                    X2=piTest$diffLL[2:5],
                    df=piTest$diffdf[2:5],
                    p=piTest$p[2:5],
                    p.bonferroni=p.adjust(p=piTest$p[2:5], method="bonferroni"),
                    p.BH=p.adjust(p=piTest$p[2:5], method="BH"))
piOut
```


