% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  man,
  floatsintext]{apa7}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic,shorthands=off]{babel}
\else
\usepackage[bidi=default,shorthands=off]{babel}
\fi
\ifLuaTeX
  \usepackage{selnolig} % disable illegal ligatures
\fi
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
% preamble.tex 

\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{caption}
\usepackage{geometry}
\usepackage{float}
\usepackage{rotating}


\captionsetup{justification=raggedright,singlelinecheck=false}

\shorttitle{SEMI-COMPLETE BY DESIGN}
\affiliation{$^{1}$Goethe University Frankfurt\\ $^{2}$Max Planck Institute for Human Development{,} Berlin{,} Germany\\ $^{3}$MSB Medical School Berlin{,} Berlin{,} Germany\\ $^{4}$Max Planck UCL Centre for Computational Psychiatry and Ageing Research{,} Berlin{,} Germany}
\usepackage{placeins}
\setlength{\textfloatsep}{6pt}
\setlength{\floatsep}{6pt}
\setlength{\intextsep}{6pt}
\raggedbottom
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={SEMi-Manuscript},
  pdfauthor={Leonie Hagitte\^{}\{1,2,3\}, Andreas M. Brandmaier\^{}\{2,3,4\}},
  pdflang={american},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{SEMi-Manuscript}
\author{Leonie Hagitte\(^{1,2,3}\), Andreas M. Brandmaier\(^{2,3,4}\)}
\date{2025-12-10}

\begin{document}
\maketitle

Psychological research frequently is concerned with latent constructs
that cannot be observed directly, such as intelligence or personality.
Because these constructs are not directly measurable, they are inferred
from manifest indicators, such as item responses or behavioral measures,
using measurement models. Consequently, the validity of substantive
conclusions depends fundamentally on the quality of the measurement
process. If a measure does not adequately represent the construct of
interest, even highest statistical rigor cannot prevent misleading
conclusions (Flake and Fried 2020; Flake et al. 2017). However, recent
work still has emphasized the need for stronger statistical methods
(Flake et al. 2017; Haig 2020; Altoè et al. 2020) as well as more
explicit theoretical foundations in psychology (Eronen and Bringmann
2021). Although these discussions often focus on issues of statistical
inference or theory construction, both ultimately depend on the quality
of measurement. Regardless of the sophistication of a statistical model
or the strength of a theoretical framework, conclusions remain limited
if the underlying measurement instrument does not function as intended.

A central aspect of measurement quality is measurement invariance (MI),
which refers to the assumption that a measurement model operates
equivalently across individuals, groups, contexts, or measurement
occasions. Under MI, the relationship between a latent construct and its
observed indicators remains stable across the conditions being compared
(Meredith 1993). This assumption is essential whenever researchers seek
to compare latent means, variances, covariances, or structural
relations, making MI a fundamental prerequisite for many correlational,
experimental, and longitudinal research designs (Putnick and Bornstein
2016).\\
Violations of MI imply that the relationship between the latent
construct and its observed indicators differs across groups or
conditions. In such cases, differences in observed responses may not
solely reflect differences in the underlying construct but may also
arise from systematic differences in item functioning or measurement
precision. Consequently, observed group differences become ambiguous, as
they may represent a combination of substantive construct differences
and measurement-related distortions (Cheung and Rensvold 2002; Meredith
1993). Without establishing MI, it is therefore impossible to determine
whether apparent differences reflect true variation in the latent
construct or artifacts of the measurement process. Consequently, MI is
not only a methodological requirement but also a prerequisite for valid
theoretical interpretation.

\subsection{Classical Measurement Invariance
Testing}\label{classical-measurement-invariance-testing}

Structural equation models (SEMs) are widely used to define latent
constructs and their relations to observed indicators. Within SEMs,
measurement models specify the relationships between latent factors and
their manifest indicators. By separating construct-related variance from
measurement error, latent variable models provide a principled basis for
evaluating measurement quality (Bollen 1989).

Classical approaches to measurement invariance testing are fundamentally
confirmatory in nature. Researchers begin by specifying a measurement
model, typically in the form of a confirmatory factor analysis (CFA),
that reflects substantive hypotheses about the construct under
investigation. A categorical grouping variable is then used to partition
the sample into two or more groups, across which measurement parameters
are constrained and compared. This most widely used approach for
evaluating measurement invariance is usually referred to as
multiple-group confirmatory factor analysis (MGCFA) (e.g. Cheung and
Rensvold 2002; Putnick and Bornstein 2016).\\
In MGCFA, measurement invariance is typically assessed through a
sequence of nested model comparisons in which equality constraints are
imposed on increasingly restrictive sets of parameters. Equality of the
models structure in both groups is called configural invariance, and can
be understood as both groups share the same factor structure
(i.e.~number of latent factors and items). Equality of factor loadings
former is referred to as metric invariance, if metric invariance is
present it indicates that all factor loading parameters are equal across
groups. Speciﬁcally, this suggests that the items of a given scale and
their underlying constructs relate to each other with the same strengths
in both groups. Scalar invariance further constrains item intercepts to
equality across groups. Establishing scalar invariance is generally
required for meaningful comparisons of latent means because it indicates
that the measurement scale has the same unit and origin across groups.
Residual or strict invariance assumes, additionally to the constraints
above, that variables residuals are equal across the groups. If the
residual variance holds, the items of the scale are measuring the latent
constructs with the same measurement error. This implies that it can
indeed be known whether diﬀerences measured by the items of the tested
scale between the groups are due to group diﬀerences in the proposed
factors or due to measurement errors (Cheung and Rensvold 2002; Widaman
and Reise 1997). The level of invariance required depends on the
substantive inference of interest. For example, scalar invariance is
generally required when comparing latent means across groups, whereas
less restrictive forms of invariance may be sufficient for other
research questions (Putnick and Bornstein 2016).

For MGCFA testing, the assessment of measurement invariance depends on
two central specifications: the measurement model itself and the
grouping variable used to represent potential sources of heterogeneity.
As a result, measurement noninvariance (MNI) can only be detected with
respect to the grouping variables included in the analysis.
Consequently, this framework is highly effective when substantive theory
naturally implies discrete group distinctions, such as treatment
conditions, demographic categories, or measurement occasions. However,
many characteristics that may influence measurement functioning vary
continuously across individuals. In such situations, representing
heterogeneity through predefined groups may require arbitrary
categorization and can obscure important variation in measurement
parameters.

\subsection{Moderated Nonlinear Factor
Analysis}\label{moderated-nonlinear-factor-analysis}

These considerations have motivated the development of approaches that
conceptualize MNI as a continuous process operating at the individual
level. One prominent example is the Moderated nonlinear factor analysis
(MNLFA). MNLFA conceptualizes heterogeneity in data as functions of
observed moderator variables. Within MNLFA, measurement parameters are
allowed to vary systematically as functions of one or more moderator
variables. Consequently, MNI is conceptualized as moderation of the
measurement model itself. If measurement parameters differ as a function
of the moderator, the measurement model no longer operates equivalently
across individuals and therefore violates measurement invariance.

Whether MI is present depends on which parameters are moderated by the
exogenous variable. The exogenous variable may influence any subset of
the parameters in the measurement model. Specifically, the means,
variances, and covariance of latent factors, as well as the item
intercepts, factor loadings, and residual variances of any manifest
indicator, may vary systematically across individuals as deterministic
functions of the exogenous variable. If moderation is limited to the
latent-factor parameters, that is, the means, variances, and covariance,
the model remains consistent with MI. In contrast, moderation of
item-level parameters, including intercepts, loadings, or residual
variances, constitutes MNI. Accordingly, a moderation function must be
specified for each parameter that is allowed to vary as a function of
the exogenous variable. Because of this conceptualization,this method
allows researchers to model gradual, nonlinear, and individual-specific
forms of MNI that are difficult to represent within traditional
multiple-group approaches.

However, this increased flexibility comes at the cost of greater
specification requirements. Like MGCFA, moderated nonlinear factor
analysis (MNLFA) constitutes a fundamentally confirmatory approach to
the assessment of MI. In practice, at least three sets of decisions are
required. First, the measurement model must be specified, including the
latent structure and the relationships between latent and observed
variables. Second, one or more moderator variables must be selected to
represent potential sources of heterogeneity in the measurement process.
Third, moderation functions must be specified for all parameters that
are allowed to vary as a function of the moderator. This includes
determining which parameters are moderated as well as specifying the
functional form of each moderation effect (e.g., linear, quadratic, or
another parametric form). Consequently, the detection of measurement
non-invariance depends not only on the choice of moderator variables but
also on the assumptions made about how these moderators influence the
model parameters.

\subsection{Structural Equation Model
Trees}\label{structural-equation-model-trees}

Structural Equation Model trees (SEM trees) were originally introduced
by Brandmaier et al.~(2013) as a method for identifying important
covariates and their interactions within SEMs. Subsequent developments
further refined the methodology and expanded its applicability (Arnold
et al., 2021; Brandmaier et al., 2016). Notably, although measurement
invariance was used as an illustrative application by Arnold et
al.~(2021), SEM Trees were not originally developed as a dedicated
method for testing measurement invariance. Because violations of MI
manifest as systematic heterogeneity in measurement parameters, SEM
trees appear conceptually well suited for detecting such effects.

SEM trees build on two paradigms: SEMs and on decision trees. Decision
trees describe recursive partitions of the covariate space, whereby the
sample is repeatedly split according to covariate values into
progressively more homogeneous subsets, in order to maximize differences
in a given outcome variable. SEM Trees combine the hierarchical
structure of the decision trees with SEMs, so that they describe
heterogeneity in data with respect to a specified SEM as targeted
outcome structure. Unlike many other tree-based methods, SEM Trees can
be applied to models that include latent variables. Despite their
conceptual suitability for this purpose, SEM trees have received
relatively little attention as tools for detecting MNI.

In contrast to MGCFA and MNLFA, SEM trees represent a more exploratory
approach to the assessment of MI. Similar to these methods, SEM trees
require the researcher to specify a measurement model a priori. Beyond
this, however, the researcher only needs to provide a set of candidate
partitioning variables that may account for heterogeneity in the data.
These variables may be numerous and can be either categorical or
continuous. In this respect, the researcher decisions required by SEM
trees closely resemble those of the traditional MGCFA framework.\\
Similar to MGCFA, researchers specify a measurement model and identify
variables that may account for heterogeneity. The crucial difference is
that MGCFA requires the researcher to select a specific grouping
variable and define its manifestation in advance, whereas SEM trees can
evaluate a large set of candidate variables simultaneously and determine
which, if any, are associated with violations of MI. Moreover, SEM trees
are able to identify nonlinear forms of moderation and data-driven
partition points for continuous variables without requiring these
relationships to be specified a priori.

Consequently, SEM trees retain the confirmatory component associated
with the measurement model while adopting an exploratory approach to the
identification of heterogeneity. This combination makes them
particularly attractive for researchers who are confident in their
measurement model but have limited theoretical guidance regarding the
sources or structure of MNI. Because the required researcher decisions
largely mirror those of the familiar MGCFA framework, the transition
from traditional MI testing to SEM trees should involve a comparatively
low methodological threshold. At the same time, SEM trees offer
substantially greater flexibility by accommodating multiple candidate
moderators, continuous and categorical partitioning variables, and
potentially nonlinear forms of heterogeneity. Although explicit
specification of moderators and moderation functions remains
advantageous when strong theoretical expectations are available, SEM
trees provide a complementary strategy for situations in which the
sources or functional forms of heterogeneity are uncertain.

\subsection{Present Study}\label{present-study}

More generally, MGCFA, MNLFA, and SEM trees can be understood as
reflecting different assumptions regarding prior knowledge about
measurement heterogeneity. MGCFA assumes that relevant sources of
heterogeneity can be represented through predefined groups. MNLFA
further assumes that researchers can identify relevant moderators and
specify how these moderators influence model parameters. SEM trees, in
contrast, place fewer assumptions on the structure of heterogeneity and
instead emphasize data-driven identification of parameter differences.
Consequently, the relative performance of these approaches is likely to
depend on the extent to which researchers possess accurate prior
knowledge about the sources and functional forms of MNI.

The present simulation study compares the performance of MNLFA and SEM
trees in detecting MNI across a range of conditions that vary in sample
size, reliability, effect magnitude, and the form of moderation. To
reflect both the methodological capabilities of the investigated
approaches and plausible empirical scenarios, the simulation
incorporates different functional forms of MNI, including linear,
quadratic, and sigmoid moderation processes. We expect SEM trees to
perform particularly well in conditions characterized by nonlinear
moderation, especially sigmoid relationships, whereas their relative
advantage should diminish in purely linear settings. In contrast, MNLFA
is expected to perform best when the functional form of moderation is
correctly specified, particularly in linear conditions, but to be more
sensitive to misspecification of the underlying moderation function. For
both methods, stronger moderation effects are expected to increase the
detectability of measurement non-invariance.

To enhance the study's relevance for empirical practice, the simulation
is grounded in a realistic applied-research perspective. Specifically,
we consider researchers whose experience with measurement invariance
testing is rooted in the traditional MGCFA framework and who seek to
adopt more flexible approaches for detecting heterogeneity. Accordingly,
MNLFA and SEM trees are evaluated not only as statistical procedures but
also as methodological alternatives for researchers moving beyond
conventional MGCFA-based invariance testing.

\section{Method}\label{method}

The simulation study was conducted in two complementary studies that
differ in the complexity of the underlying latent-variable model.
``Study 1'' focuses on MNI within a single-factor framework and
investigates how the performance of MNLFA and SEM trees is affected by
characteristics of the moderation process, including the number of
moderated parameters, the strength and functional form of moderation,
sample size, and indicator reliability. Particular emphasis is placed on
conditions in which the analytical model may incorrectly represent the
true moderation function. ``Study 2'' extends this framework to a
two-factor model. In addition to measurement-level noninvariance,
heterogeneity may also occur in relations among latent variables through
moderation of the latent covariance. Furthermore, ``Study 2''
incorporates structural-model misspecifications, allowing the robustness
of both methods to be evaluated under more complex latent-variable
structures.

\subsection{Data-Generation Mechanism}\label{data-generation-mechanism}

Across all simulation conditions, MNI was generated through moderation
of model parameters by an observed continuous covariate. A raw moderator
M was sampled from a uniform distribution on the interval \([-1,1]\) and
transformed into an effective moderator \(Z=h(M)\). The transformation
\(h(\cdot)\) determined the functional form of heterogeneity and varied
across simulation conditions. Importantly, nonlinearity was introduced
exclusively through the transformation \(Z=h(M)\), while moderation
effects remained linear in the transformed moderator. Model parameters
subject to moderation were specified as deterministic functions of
\(Z\), such that measurement properties changed systematically across
individuals.

\subsubsection{Functional Form of
Moderation}\label{functional-form-of-moderation}

The effective moderator \(Z = h(M)\) was generated systematically using
the following DGM functions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Linear\\
\item
  Quadratic\\
\item
  Sigmoid\\
\item
  Noise/null moderation, \(h(M)=0\)
\end{enumerate}

These functions were selected to represent qualitatively different forms
of heterogeneity. Linear moderation reflects the functional form most
commonly assumed in modeling applications,implying a constant change in
measurement parameters across the moderator range.\\
In contrast, quadratic moderation captures situations in which
measurement differences are strongest at extreme values of a
characteristic, such as when respondents at very low and very high
levels of a trait interpret items differently than individuals near the
center of the distribution. Sigmoid moderation reflects threshold-like
processes in which measurement parameters change little across most of
the moderator range but shift more rapidly once a critical region is
reached, for example when item functioning remains stable until a
particular level of symptom severity, cognitive ability, or
developmental maturation is exceeded. The null condition \((h(M)=0)\)
served as a measurement-invariant baseline and was used to evaluate Type
I error rates.

\subsubsection{Strength of Moderation}\label{strength-of-moderation}

In each case where parameters are subjected to moderation, the strength
of the moderation \(\delta\) is varied:

\(\delta_{\lambda}\) is evaluated at the following discrete levels:

\[
\delta_{\lambda} \in \{-0.3,\,-0.2,\,0.2,\,0.3\}.
\]

\[
\delta_{\nu} \in \{-1.0,\,-0.5,\,0.5,\,1.0\} \quad \text{for } i \in \{1,2\},
    \qquad
    \delta_{\nu} = 0 \quad \text{for } i \in \{3,4\},
\]

\subsubsection{Model Misspecification of Moderator
Form}\label{model-misspecification-of-moderator-form}

The DGM always defines the effective moderator through \(Z=h(M)\). In
MNLFA, the analytical model is fitted using prespecified parametric
forms, namely linear and quadratic moderation. Consequently, MNLFA is
correctly specified when the fitted analytical form corresponds to the
DGM transformation and misspecified otherwise. SEM trees do not impose a
parametric functional form, but they must identify the correct moderator
variable(s) from the candidate set. The simulation assumes that the
moderators are always perfectly measured. So this really reflects an
ideal scenario and should be kept in mind, when interpreting the
results. We expect that one should see a similar picture, but with
less-than-perfect reliabilities and lower power in general, should one
add variation to the reliability of the moderators.

\subsubsection{Performance Criteria}\label{performance-criteria}

The primary estimands in this simulation study pertain to the detection
and accurate estimation of MNI due to moderation effects on measurement
parameters. Specifically, we assess whether the estimation methods
compared can correctly identify the presence or absence of moderation in
factor loadings, intercepts, and residual variances, reflecting the true
data-generating structure. Consequently, key simulation outcomes include
Type I error rates and statistical power associated with detecting such
moderation effects, as well as bias and variability in the recovered
parameter estimates. Secondary estimands include model-level indicators
of fit (e.g., Kullback--Leibler divergence, AIC), which serve as
auxiliary diagnostics for methods that yield such metrics (i.e., MNLFA).
These are not to the same extent applicable to tree-based approaches and
are thus interpreted as method-specific targets rather than general
evaluative criteria.

For the parametric MNLFA, we will extract estimated moderated parameters
(e.g., factor loadings, intercepts), their associated standard errors,
and two-sided p-values for testing the null hypothesis of no moderation
effect. Additionally, model fit indices such as Kullback--Leibler
divergence, AIC and the LRT-statistics will be recorded. For the
parameter estimates and fit indices, we will summarize their
distributions across simulation replications by reporting key quantiles
(2.5th, 50th, and 97.5th percentiles). The null hypothesis will be
rejected if the associated p-value is less than the conventional
significance level of 0.05. In contrast, non-parametric tree-based
approaches do not yield direct parameter estimates with standard errors.
We use score-based SEM-trees (Arnold et al. 2021). Their detection of
MI/ non-invariance is operationalized via the presence of at least one
split in the correct condition at the correct respective level of focus
parameters. Those splits, introducing a subgroup, will be evaluated at
the significance level: 0.05. Additionally, a Bonferroni correction to
account for multiple comparisons at each node. The trees will be allowed
to split on the focus parameters relevant for the invariance testing
(the means of the intercepts, factor loadings and the latent covariance
in study 2). Candidate subgroup structures are assessed by testing
whether they produce statistically significant differences in the
specified model parameters across groups. Differences in factor
variances between subgroups are permitted in the model, but they are not
used as criteria for selecting or evaluating covariates. Furthermore,
pre-pruning will be applied, limiting the maximum depth of the tree to 3
and min.\textbf{N} is set to 50. Because the MNLFA also directly
considers the form of the moderation, which the SEM trees do not, we
further want to use SEM forests for generating partial dependence plots
to get an approximation of the functional form of the moderating
variables.

\subsubsection{Replications}\label{replications}

We employ a full grid search, in which parameter values are sampled from
pre-specified discrete values. To ensure computational feasibility
however, we followed an iterative approach whereby a initial pilot
simulations \(n=100\) was conducted to amongst others, estimate the
needed simulation time. These estimates then informed the total number
of XX replications.

\subsubsection{Software and
Implementation}\label{software-and-implementation}

All simulations and analyses will be conducted in R (R Core Team 2025).
The MNLFA-style models will be implemented directly in
\texttt{OpenMx}(Boker et al. 2026), using raw-data maximum likelihood
estimation and definition-variable-style moderation through OpenMx
algebra objects. SEM tree analyses will be performed with the
\texttt{semtree} (Brandmaier et al. 2026) package using OpenMx RAM
models. Simulation design construction, data handling, aggregation, and
visualization will rely on \texttt{dplyr}(Wickham et al. 2026),
\texttt{tidyr} (Wickham et al. 2025), \texttt{tibble} (Müller and
Wickham 2026), \texttt{purrr} (Wickham and Henry 2026), \texttt{furrr}
(Vaughan and Dancho 2022), \texttt{future} (Bengtsson 2021),
\texttt{parallel} (R Core Team 2025), and \texttt{ggplot2} (Wickham
2016). The simulation study will be conducted in R (R Core Team 2025).

\subsubsection{General Analysis
Procedures}\label{general-analysis-procedures}

\subsection{Study 1}\label{study-1}

\subsubsection{Population Models}\label{population-models}

\subsubsection{Simulation Conditions}\label{simulation-conditions}

Focuses on functional form misspecification and data-related conditions:

\begin{itemize}
\tightlist
\item
  \textbf{Sample sizes:} \(N \in \{300, 500, 700, 1,000\}\)
\item
  \textbf{Intercepts:} Intercept baselines are set to 1 across
  simulation conditions.\\
\item
  \textbf{Indicator reliabilities:} \emph{Low} (.6), \emph{moderate}
  (.7), \emph{moderate to high} (.8), or \emph{high} (.95).
\end{itemize}

\subsubsection{Results: Study 1}\label{results-study-1}

\subsection{Study 2}\label{study-2}

\subsubsection{Population Models}\label{population-models-1}

\subsubsection{Simulation Conditions}\label{simulation-conditions-1}

Builds on Study 1 and introduces additional model misspecification in
the factor structure:

\begin{itemize}
\tightlist
\item
  \textbf{Structural model misspecifications:} (between-subjects factor)

  \begin{itemize}
  \tightlist
  \item
    \textbf{1. No misspecification:} Factor structure matches the
    analysis model.
  \item
    \textbf{2. Cross-loadings:} Items \(y_5\), \(y_6\) additionally load
    on a second latent factor (See Figure 3).\\
  \end{itemize}
\item
  \textbf{Latent Covariance:} The latent covariance is moderated by
  \(Z\). The baseline variance is set to 1 and moderation strength is
  selected from:
\end{itemize}

\[
\delta_{\sigma_\eta} \in \{-0.3,\,-0.2,\,0.2,\,0.3\}.
\]

\subsubsection{Results: Study 2}\label{results-study-2}

\section{Discussion}\label{discussion}

\subsection{Practical Recommendations}\label{practical-recommendations}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\protect\phantomsection\label{refs}
\begin{CSLReferences}{1}{1}
\bibitem[\citeproctext]{ref-Altoe2020}
Altoè, Gianmarco, Giulia Bertoldo, Claudio Zandonella Callegher, et al.
2020. {``Enhancing Statistical Inference in Psychological Research via
Prospective and Retrospective Design Analysis.''} \emph{Frontiers in
Psychology}, advance online publication.
\url{https://doi.org/10.3389/fpsyg.2019.02893}.

\bibitem[\citeproctext]{ref-Arnold2021}
Arnold, M., M. C. Voelkle, and A. M. Brandmaier. 2021. {``Score-Guided
Structural Equation Model Trees.''} \emph{Frontiers in Psychology} 11:
3913. \url{https://doi.org/10.3389/fpsyg.2020.564403}.

\bibitem[\citeproctext]{ref-future2021}
Bengtsson, Henrik. 2021. {``A Unifying Framework for Parallel and
Distributed Processing in r Using Futures.''} \emph{The R Journal} 13
(2): 208--27. \url{https://doi.org/10.32614/RJ-2021-048}.

\bibitem[\citeproctext]{ref-Boker2026}
Boker, Steven M., Michael C. Neale, Hermine H. Maes, et al. 2026.
\emph{OpenMx: Extended Structural Equation Modelling}.
\url{https://doi.org/10.32614/CRAN.package.OpenMx}.

\bibitem[\citeproctext]{ref-Bollen1989}
Bollen, K. A. 1989. \emph{Structural Equations with Latent Variables}.
John Wiley \& Sons. \url{https://doi.org/10.1002/9781118619179}.

\bibitem[\citeproctext]{ref-semtree2026}
Brandmaier, A. M., John J. Prindle, Manuel Arnold, and Caspar J. Van
Lissa. 2026. \emph{Semtree: Recursive Partitioning for Structural
Equation Models}. \url{https://github.com/brandmaier/semtree}.

\bibitem[\citeproctext]{ref-Cheung2002}
Cheung, Gordon W., and Roger B. Rensvold. 2002. {``Evaluating
Goodness-of-Fit Indexes for Testing Measurement Invariance.''}
\emph{Structural Equation Modeling: A Multidisciplinary Journal} 9 (2):
233--55. \url{https://doi.org/10.1207/S15328007SEM0902/_5}.

\bibitem[\citeproctext]{ref-Eronen2021}
Eronen, Markus I., and Laura F. Bringmann. 2021. {``The Theory Crisis in
Psychology: How to Move Forward.''} \emph{Perspectives on Psychological
Science} 16 (4): 779--88.
\url{https://doi.org/10.1177/1745691620970586}.

\bibitem[\citeproctext]{ref-Flake2020}
Flake, Jessica K., and Eiko I. Fried. 2020. {``Measurement
Schmeasurement: Questionable Measurement Practices and How to Avoid
Them.''} \emph{Advances in Methods and Practices in Psychological
Science} 3 (4): 456--65. \url{https://doi.org/10.1177/2515245920952393}.

\bibitem[\citeproctext]{ref-Flake2017}
Flake, Jessica K., Jolynn Pek, and Eric Hehman. 2017. {``Construct
Validation in Social and Personality Research: Current Practice and
Recommendations.''} \emph{Social Psychological and Personality Science}
8 (4): 370--78. \url{https://doi.org/10.1177/1948550617693063}.

\bibitem[\citeproctext]{ref-Haig2020}
Haig, Brian D. 2020. {``What Can Psychology's Statistics Reformers Learn
from the Error-Statistical Perspective?''} \emph{Methods in Psychology}
2: 100020.
https://doi.org/\url{https://doi.org/10.1016/j.metip.2020.100020}.

\bibitem[\citeproctext]{ref-Meredith1993}
Meredith, William. 1993. {``Measurement {Invariance}, {Factor Analysis}
and {Factorial Invariance}.''} \emph{Psychometrika} 58 (4): 525--43.
\url{https://doi.org/10.1007/BF02294825}.

\bibitem[\citeproctext]{ref-tibble2026}
Müller, Kirill, and Hadley Wickham. 2026. \emph{Tibble: Simple Data
Frames}. \url{https://doi.org/10.32614/CRAN.package.tibble}.

\bibitem[\citeproctext]{ref-Putnick2016}
Putnick, Diane L., and Marc H. Bornstein. 2016. {``Measurement
Invariance Conventions and Reporting: The State of the Art and Future
Directions for Psychological Research.''} \emph{Developmental Review}
41: 71--90.
https://doi.org/\url{https://doi.org/10.1016/j.dr.2016.06.004}.

\bibitem[\citeproctext]{ref-Rteam2025}
R Core Team. 2025. \emph{R: A Language and Environment for Statistical
Computing}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\bibitem[\citeproctext]{ref-furrr2022}
Vaughan, Davis, and Matt Dancho. 2022. \emph{Furrr: Apply Mapping
Functions in Parallel Using Futures}.
\url{https://doi.org/10.32614/CRAN.package.furrr}.

\bibitem[\citeproctext]{ref-ggplot2016}
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
Analysis}. Springer-Verlag New York.
\url{https://ggplot2.tidyverse.org}.

\bibitem[\citeproctext]{ref-dplyr2026}
Wickham, Hadley, Romain François, Lionel Henry, Kirill Müller, and Davis
Vaughan. 2026. \emph{Dplyr: A Grammar of Data Manipulation}.
\url{https://doi.org/10.32614/CRAN.package.dplyr}.

\bibitem[\citeproctext]{ref-purrr2026}
Wickham, Hadley, and Lionel Henry. 2026. \emph{Purrr: Functional
Programming Tools}. \url{https://doi.org/10.32614/CRAN.package.purrr}.

\bibitem[\citeproctext]{ref-tidyr2025}
Wickham, Hadley, Davis Vaughan, and Maximilian Girlich. 2025.
\emph{Tidyr: Tidy Messy Data}.
\url{https://doi.org/10.32614/CRAN.package.tidyr}.

\bibitem[\citeproctext]{ref-Widaman1997}
Widaman, Keith F., and Steven P. Reise. 1997. {``Exploring the
Measurement Invariance of Psychological Instruments: {Applications} in
the Substance Use Domain.''} In \emph{The Science of Prevention:
{Methodological} Advances from Alcohol and Substance Abuse Research.}
American Psychological Association.
\url{https://doi.org/10.1037/10222-009}.

\end{CSLReferences}

\end{document}
