---
title: "My Quarto Shiny Presentation"
format:
  revealjs:
    theme: white
    css: styles.css
    scrollable: true
    slide-number: c/t
server: shiny
---

```{r}
#| label: setup
#| echo: false

library(shiny)
library(dplyr)
library(ggplot2)
library(ggalluvial)
library(stringr)
library(scales)

results2 <- readRDS("results2.rds")

to_logical <- function(x) {
  if (is.logical(x)) return(x)
  if (is.numeric(x)) return(as.logical(x))
  x <- as.character(x)
  case_when(
    x %in% c("TRUE", "True", "true", "1") ~ TRUE,
    x %in% c("FALSE", "False", "false", "0") ~ FALSE,
    TRUE ~ NA
  )
}

alpha <- 0.05

tree_flow_dat <- results2 %>%
  filter(moderator != "noise") %>%
  mutate(
    truth = case_when(
      true_metric_noninvariance == TRUE & true_scalar_noninvariance == TRUE ~ "Metric & Scalar",
      true_metric_noninvariance == TRUE & true_scalar_noninvariance == FALSE ~ "Metric only",
      true_metric_noninvariance == FALSE & true_scalar_noninvariance == TRUE ~ "Scalar only",
      true_metric_noninvariance == FALSE & true_scalar_noninvariance == FALSE ~ "Invariant",
      TRUE ~ NA_character_
    ),
    
    metric_tree_decision = case_when(
      tree_metric_reject_bonf == TRUE ~ "Metric tree rejected",
      tree_metric_reject_bonf == FALSE ~ "Metric tree not rejected",
      is.na(tree_metric_reject_bonf) ~ "Metric tree missing"
    ),
    
    scalar_tree_decision = case_when(
      tree_metric_reject_bonf == TRUE ~ "Scalar tree not attempted",
      tree_metric_reject_bonf == FALSE & tree_scalar_reject_bonf == TRUE ~ "Scalar tree rejected",
      tree_metric_reject_bonf == FALSE & tree_scalar_reject_bonf == FALSE ~ "Scalar tree not rejected",
      is.na(tree_scalar_reject_bonf) ~ "Scalar tree missing",
      TRUE ~ "Other"
    ),
    
    final_tree_outcome = case_when(
      true_metric_noninvariance == TRUE &
        tree_metric_reject_bonf == TRUE &
        tree_metric_correct_split == TRUE ~ "Correct metric detection",
      
      true_scalar_noninvariance == TRUE &
        tree_metric_reject_bonf == FALSE &
        tree_scalar_reject_bonf == TRUE &
        tree_scalar_correct_split == TRUE ~ "Correct scalar detection",
      
      true_metric_noninvariance == FALSE &
        tree_metric_reject_bonf == TRUE ~ "Metric false positive",
      
      true_scalar_noninvariance == FALSE &
        tree_metric_reject_bonf == FALSE &
        tree_scalar_reject_bonf == TRUE ~ "Scalar false positive",
      
      true_any_noninvariance == TRUE &
        (tree_metric_reject_bonf == FALSE | is.na(tree_metric_reject_bonf)) &
        (tree_scalar_reject_bonf == FALSE | is.na(tree_scalar_reject_bonf)) ~ "Missed noninvariance",
      
      true_any_noninvariance == FALSE &
        tree_metric_reject_bonf == FALSE &
        tree_scalar_reject_bonf == FALSE ~ "Correct invariance decision",
      
      TRUE ~ "Missing"
    )
  ) %>%
  count(truth, metric_tree_decision, scalar_tree_decision, final_tree_outcome, name = "n") %>%
  mutate(prop = n / sum(n))


#------------------------------------------------------------------------------
truth_cols <- c(
  "Metric & Scalar" = "#b38711",
  "Metric only"     = "#af4f2f",
  "Scalar only"     = "#732f30",
  "Invariant"       = "#59385c"
)
axis_levels <- c(
  "truth",
  "metric_tree_decision",
  "scalar_tree_decision",
  "final_tree_outcome"
)

axis_labels <- c(
  truth = "Truth",
  metric_tree_decision = "Metric tree",
  scalar_tree_decision = "Scalar tree",
  final_tree_outcome = "Final outcome"
)

truth_order <- c(
  "Metric & Scalar",
  "Metric only",
  "Scalar only",
  "Invariant"
)
```

## 
:::{.highlight-box}
Do students with higher socioeconomic status report higher academic motivation?
:::

::: {.incremental .small}
- Socioeconomic status is **continuous**.
- Measurement invariance is typically evaluated using **groups**.
- Researchers therefore create categories such as **low**, **medium**, and **high** SES.
- But these categories are **arbitrary**.
:::

::: {.fragment}
<div class="takeaway">
Can we evaluate measurement invariance without categorizing SES?
</div>
:::

## Why this is a measurement problem

Latent variables are not observed directly.

They are inferred from item responses.

$$
x_{ij} = \nu_j + \lambda_j \eta_i + \varepsilon_{ij}
$$

::: {style="font-size:0.5em;"}

- $\eta_i$ is the latent trait of person $i$.
- $\lambda_j$ links item $j$ to the latent trait.
- $\nu_j$ captures the expected item response when $\eta_i = 0$.
- Meaningful comparisons require that these measurement parameters are stable.

:::

::: {.fragment}

<div class="takeaway">
Before comparing latent motivation across SES, we must ask whether the items measure motivation equivalently.
</div>

:::

##

<div class="student-diagram">

<div class="student-row">
  <div class="person"><strong>Student A</strong><br><span>Low SES</span></div>
  <div class="person"><strong>Student B</strong><br><span>High SES</span></div>
</div>

<div class="student-row">
  <div class="scale">📋 Motivation scale</div>
  <div class="scale">📋 Motivation scale</div>
</div>

<div class="student-row fragment">
  <div class="arrow">↓</div>
  <div class="arrow">↓</div>
</div>

<div class="student-row fragment">
  <div class="latent">Latent motivation</div>
  <div class="latent">Latent motivation</div>
</div>

</div>


## Shiny slide {.compact-slide}

:::: {.columns}
::: {.column width="90%"}

```{r}

div(
  class = "compact-shiny-output",
  
  verbatimTextOutput(
    "step_label",
    placeholder = FALSE
  ),
  
  div(
    style = "overflow: hidden;",
    class = "compact-plot",
    plotOutput(
      "tree_flow_plot",
      width = "100%",
      height = "650px"
    )
  )
)

```

:::
::: {.column width="10%"}

```{r}
div(
   style = "margin-top: 500px; margin-left: 50px;",
  actionButton(
  "next_step",
  HTML("&#8250;"),
  class = "subtle-button",
  style = "
    color: rgba(120,120,120,0.25);
    background-color: transparent;
    border: none;
    box-shadow: none;
    font-size: 22px;
    width: 36px;
    height: 36px;
    padding: 0;
    border-radius: 50%;
  "
)
)
```

:::
::::

```{r}
#| context: server

step <- reactiveVal(1)

observeEvent(input$next_step, {
  next_value <- step() + 1
  
  if (next_value > length(truth_order)) {
    next_value <- 1
  }
  
  step(next_value)
})

output$step_label <- renderText({
  paste(
    "Currently shown:",
    paste(truth_order[seq_len(step())], collapse = ", ")
  )
})

output$tree_flow_plot <- renderPlot({
  plot_dat <- tree_flow_dat %>%
    filter(truth %in% truth_order[seq_len(step())])

  # your full ggplot code here
  tree_flow_plot_perc <- plot_dat %>%
      mutate(
        alluvium_id = row_number(),
        flow_truth = truth
      )
    
    tree_lodes <- to_lodes_form(
      tree_flow_plot_perc,
      axes = all_of(axis_levels),
      id = "alluvium_id",
      key = "axis",
      value = "stratum"
    ) %>%
      mutate(
        axis = factor(axis, levels = axis_levels, labels = axis_labels)
      )
    
    stratum_props <- tree_lodes %>%
      group_by(axis, stratum) %>%
      summarise(n = sum(n), .groups = "drop") %>%
      group_by(axis) %>%
      mutate(
        pct = n / sum(n),
        pct_label = scales::percent(pct, accuracy = 0.1)
      ) %>%
      ungroup()
    
    ggplot(
      tree_lodes,
      aes(
        x = axis,
        stratum = stratum,
        alluvium = alluvium_id,
        y = n
      )
    ) +
      geom_alluvium(
        aes(fill = flow_truth),
        alpha = 0.60,
        width = 0.24,
        knot.pos = 0.45,
        curve_type = "sigmoid",
        colour = NA
      ) +
      geom_stratum(
        data = tree_lodes %>% filter(axis != "Truth"),
        fill = "#dcdcdc",
        colour = "white",
        linewidth = 1,
        width = 0.10,
        alpha = 1
      ) +
      geom_stratum(
        data = tree_lodes %>% filter(axis == "Truth"),
        aes(fill = stratum),
        colour = "white",
        linewidth = 1,
        width = 0.24,
        alpha = 1
      ) +
      geom_text(
        data = stratum_props %>% filter(axis == "Truth"),
        inherit.aes = FALSE,
        stat = "stratum",
        aes(
          x = axis,
          stratum = stratum,
          y = n,
          label = paste0(
            str_wrap(stratum, width = 14),
            "\n",
            pct_label
          )
        ),
        hjust = 0.5,
        size = 6.5,
        fontface = "bold",
        lineheight = 0.9,
        colour = "black"
      ) +
      geom_text(
        data = stratum_props %>% filter(axis != "Truth"),
        inherit.aes = FALSE,
        stat = "stratum",
        aes(
          x = axis,
          stratum = stratum,
          y = n,
          label = pct_label
        ),
        size = 6.0,
        fontface = "bold",
        colour = "grey25"
      ) +
      geom_text(
        data = stratum_props %>% filter(axis != "Truth"),
        inherit.aes = FALSE,
        stat = "stratum",
        aes(
          x = axis,
          stratum = stratum,
          y = n,
          label = str_wrap(stratum, width = 20)
        ),
        hjust = 0,
        nudge_x = 0.08,
        size = 6.0,
        lineheight = 0.9,
        colour = "black"
      ) +
      scale_x_discrete(
        expand = expansion(mult = c(0.00, 0.00))
      ) +
      scale_fill_manual(
        values = truth_cols,
        breaks = truth_order[seq_len(step())]
      ) +
      labs(
        x = NULL,
        y = NULL,
        fill = NULL
      ) +
      coord_cartesian(clip = "off") +
      theme_minimal(base_size = 24) +
      theme(
        panel.grid = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_text(size = 24, face = "bold"),
        legend.position = "bottom",
        plot.margin = margin(10, 120, 10, 10)
      )
  })
```